---
title: "lab5-Chen"
author: "Weixuan Chen"
date: "3/1/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data preparation

```{r}
d1 <- read.csv("http://faraway.neu.edu/biostats/lab5_dataset1.csv")
d2 <- read.csv("http://faraway.neu.edu/biostats/lab5_dataset2.csv")
```

```{r}
d1
```
## Task1

## 1
H0: There is no differemce in means among all groups
H1: Not H0 (there is difference)

## 2

```{r}
control_mort <- subset(d1, treatment == 'control')
t1_mort <- subset(d1, treatment == 't1')
t2_mort <- subset(d1, treatment == 't2')
t3_mort <- subset(d1, treatment == 't3')
t4_mort <- subset(d1, treatment == 't4')
hist(control_mort$mortality)
hist(t1_mort$mortality)
hist(t2_mort$mortality)
hist(t3_mort$mortality)
hist(t4_mort$mortality)
```
No, they do not seem normally distributed.

## 3

```{r}
t1_mort_trans <- log10(t1_mort$mortality)
t2_mort_trans <- log10(t2_mort$mortality)
```

```{r}
par(mfrow=c(2,2))
hist(t1_mort_trans)
hist(t2_mort_trans)
```
Now they seem to be normally distributed. The transformation is necessary.
I would use transformed data.

## 4

```{r}
d1['mortality_transformed'] <- log10(d1$mortality)
mod.aov.mort <- aov(mortality ~ treatment, data = d1)
summary(mod.aov.mort)
```


## 5
From te table, we can see the p-value are significant. We can reject the null hypothesis. There is difference between means among groups.

## 6
Because Tukey's HSD controls the family-wise error while t-tests don't.

## 7

```{r}
thsd <- TukeyHSD(mod.aov.mort)
thsd
```

## 8

```{r}
plot(thsd)
```
t3-t2, and t4-t3 are different from the mean

## 9
```{r}
##install.packages("multcompView")
require(multcompView)
```


## 10
```{r}
mortality_lables <-multcompLetters(thsd$treatment [, "p adj"])$Letters


mortality_lables_order <- mortality_lables[order(names(mortality_lables))]

print (mortality_lables_order)
```

## 11

```{r}
mean.mortality <- aggregate(mortality ~ treatment, data = d1, FUN = mean)

stderr.mortality <- aggregate(mortality ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)),data = d1)

ci.upp <- mean.mortality$mortality + 1.96 * stderr.mortality$mortality

ci.low <- mean.mortality$mortality - 1.96 * stderr.mortality$mortality

bp <- barplot(mean.mortality$mortality, names = mean.mortality$treatment,
              main = "Mortality Rate for Treatments",
              ylim = c(0, max(ci.upp + 1.9)), ylab = "Mortality")

arrows(y0 = ci.low, y1 = ci.upp, x0 = bp, x1 = bp, angle = 90,
       code = 3, length = 0.1)

text(x = bp, y = mean.mortality$mortality + stderr.mortality$mortality,
     mortality_lables, pos = 3)
```

## 12
t4 seems most effective in increasing pest mortality because it has the hightest mortality rate and highest reaching error bar

## Task 2

## 1

```{r}
mortality_anova2 = aov(log10(mortality) ~ treatment, d2)

summary(mortality_anova2)
```
ANOVA table suggests that the mean mortality is different accross the treatments

## 2

```{r}
mortality_tukey2 = TukeyHSD(mortality_anova2)

print(mortality_tukey2)
```

## 3

```{r}
library(multcompView)

mortality_lables2 <-multcompLetters(mortality_tukey2$treatment [, "p adj"])$Letters

mortality_lables_order2 <- mortality_lables2[order(names(mortality_lables2))]

print (mortality_lables_order2)
```

```{r}
mean.mortality2 <- aggregate(mortality ~ treatment, data = d2, FUN = mean)

stderr.mortality2 <- aggregate(mortality ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)), data = d2)

ci.upp2 <- mean.mortality2$mortality + 1.96 * stderr.mortality2$mortality

ci.low2 <- mean.mortality2$mortality - 1.96 * stderr.mortality2$mortality

bp2 <- barplot(mean.mortality2$mortality, names = mean.mortality2$treatment,
              main = "Mortality Rate for Treatments",
              ylim = c(0, max(ci.upp2 + 1.9)), ylab = "Mortality")

arrows(y0 = ci.low2, y1 = ci.upp2, x0 = bp2, x1 = bp2, angle = 90,
       code = 3, length = 0.1)

text(x = bp2, y = mean.mortality2$mortality + stderr.mortality2$mortality,
     mortality_lables2, pos = 3)
```

## 4
t2-control, t4-t1,t3-t2, and t4-t2 are all different from the mean
t4 has the lowest mortality rate for the enemy of the pest

## 5
t4 is the best treatment because it has the highest mortality rate for the pest, and lowest mortality rate for the enemy of the pest.