---
title: "lab3-Chen"
author: "Weixuan Chen"
date: "2/15/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

## Task1

## 1
We can have a sample size of 100 seeds for both GMO and wild type, and plant them in the same field. Here, we control the environment they will grow and the only difference is the type of themselves. And we can count the number of germinated seeds in each group. And, for the purpose of repliaction, we can repeat the experiment for 100 times, or we have 100 trials going on simutaneously. 

## 2

```{r}
# Download the data
d1 <- read.csv("http://faraway.neu.edu/biostats/lab3_dataset1.csv")
head(d1)
```

## 3

Plot for gmo.germinated
```{r}
mean_gmo_germ = mean(d1$gmo.germinated)
upper_gmo_germ = mean_gmo_germ + 1.96*sd(d1$gmo.germinated)/sqrt(length(d1$gmo.germinated))
lower_gmo_germ = mean_gmo_germ - 1.96*sd(d1$gmo.germinated)/sqrt(length(d1$gmo.germinated))

mean_gmo_notgerm = mean(d1$gmo.notgerminated)
upper_gmo_notgerm = mean_gmo_notgerm + 1.96*sd(d1$gmo.notgerminated)/sqrt(length(d1$gmo.notgerminated))
lower_gmo_notgerm = mean_gmo_notgerm - 1.96*sd(d1$gmo.notgerminated)/sqrt(length(d1$gmo.notgerminated))

mean_wild_germ = mean(d1$wild.germinated)
upper_wild_germ = mean_wild_germ + 1.96*sd(d1$wild.germinated)/sqrt(length(d1$wild.germinated))
lower_wild_germ = mean_wild_germ - 1.96*sd(d1$wild.germinated)/sqrt(length(d1$wild.germinated))

mean_wild_notgerm = mean(d1$wild.notgerminated)
upper_wild_notgerm = mean_wild_notgerm + 1.96*sd(d1$wild.notgerminated)/sqrt(length(d1$wild.notgerminated))
lower_wild_notgerm = mean_wild_notgerm - 1.96*sd(d1$wild.notgerminated)/sqrt(length(d1$wild.notgerminated))


seeds_mean <- c(mean_gmo_germ, mean_wild_germ, mean_gmo_notgerm, mean_wild_notgerm)
seeds_lower_ci <- c(lower_gmo_germ, lower_wild_germ, lower_gmo_notgerm, lower_wild_notgerm)
seeds_upper_ci <- c(upper_gmo_germ, upper_wild_germ, upper_gmo_notgerm, upper_wild_notgerm)
seeds_index <- c('gmo.germinated', 'wild.germinated', 'gmo.notgerminated', 'wild.notgerminated')

df <- data.frame(seeds_index, seeds_mean, seeds_lower_ci, seeds_upper_ci)

df
```

## 4

```{r}
ggplot(data = df) + 
  geom_bar(aes(x=seeds_index, y=seeds_mean, fill = seeds_index), stat = "identity") + 
  geom_errorbar(aes(x=seeds_index, ymin = seeds_lower_ci, ymax = seeds_upper_ci), width = 0.4, color = "red", size = 1) + 
  geom_text(aes(x=seeds_index, y = seeds_lower_ci, label = round(seeds_lower_ci, 1)), size = 2, vjust = 1) +
  geom_text(aes(x=seeds_index, y = seeds_upper_ci, label = round(seeds_upper_ci, 1)), size = 2, vjust = -1) +
  theme_classic() + 
  labs(title = "Confidence Interval for d1") + 
  labs(x="Groups", y = "Mean Number of Seeds")
```

## 5
By the plot, we can see the mean of germinated GMO outperformed the wild type. It is definitely worth toiling in the lab.

## 6
H0: GMO seeds and wild seeds are equally likely to germinated
H1: Not H0

## 7

```{r}
germinated <- c(sum(d1$gmo.germinated), sum(d1$wild.germinated))

notgerminated <- c(sum(d1$gmo.notgerminated), sum(d1$wild.notgerminated))

fisher_index <- c('GMO', 'Wild')

fisher_df <- data.frame(germinated, notgerminated)
rownames(fisher_df) <- fisher_index

fisher_df
```

```{r}
fisher_tst <- fisher.test(fisher_df)
fisher_tst
```

## 8
We can see the p-value is 2.2e-16, which is so small that we should reject the null hypothesis whichever the significance level we choose. So there is evidence that seed germination is different in GMO and Wild groups.


## Task 2

```{r}
 # Download the data
d2 <- read.csv("http://faraway.neu.edu/biostats/lab3_dataset2.csv")
head(d2)
```

## 1
H0: two variables are independent (GMO and disease have no association)
H1: Not H0

## 2
```{r}
have_disease <- c(sum(d2$gmo.disease), sum(d2$nogmo.disease))

have_no_disease <- c(sum(d2$gmo.nodisease), sum(d2$nogmo.nodisease))

disease_index <- c('GMO', 'Not GMO')

disease_df <- data.frame(have_disease, have_no_disease)
rownames(disease_df) <- disease_index

disease_df
```

```{r}
disease_tst <- fisher.test(disease_df)
disease_tst
```

## 3
We can see the p-value is 0.05242, which is greater than 0.05 level. So we say that we don't have enough evidence to show that these two groups have no association at 0.05 level.

## 4

```{r}
pvals <- numeric(NROW(d2)) 
for (i in 1:NROW(d2)) {
  have_disease <- c(d2$gmo.disease[i], d2$nogmo.disease[i])
  have_no_disease <- c(d2$gmo.nodisease[i], d2$nogmo.nodisease[i])
  disease_index <- c('GMO', 'Not GMO')
  disease_df <- data.frame(have_disease, have_no_disease)
  rownames(disease_df) <- disease_index
  disease_tst <- fisher.test(disease_df)
  pvals[i] = disease_tst$p.value
}

df_pvals_countries <- data.frame(pvals)
rownames(df_pvals_countries) <- d2$countries
df_pvals_countries
```

## 5
Except for India, there is evidence of an association between GMO and disease at 0.05 significance level.