---
title: "proj1-WeixuanChen"
author: "Weixuan Chen"
date: "2/16/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

## 1
By Bayes Theorem,p(f2|d) = p(d|f2)*p(f2)/p(d). p(d|f2) = 0.7, p(f2) = 0.6. And by the law of total probability, p(d) = p(d|f2)p(f2) + p(d|not f2)p(not f2) = 0.7 x 0.6 + 0.1 x 0.4, so we can finalize the result in the code below.

```{r}
pDF2 = 0.7
pF2 =0.6
pDNF2 = 0.1
pD = pDF2 * pF2 + pDNF2 * (1 - pF2)
pF2D = pDF2*pF2/pD
pF2D
```

## 2

```{r}
p.F2 <- seq(0.1, 0.9, length.out = 100)
p.F2.D.1 <- c()
for (i in 1:100){
    pD.F2 = 0.7
    pD.NF2 = 0.1
    p.D = pD.F2 * p.F2[i] + pD.NF2 * (1 - p.F2[i])
    pF2.D = pD.F2*p.F2[i]/p.D
    p.F2.D.1 <- append(p.F2.D.1,pF2.D)
}

p.F2.D.1
```

```{r}
p.D.not.F2 <- seq(0.01, 0.5, length.out = 100)
p.F2.D.2 <- c()
for (i in 1:100){
    pD.F2 = 0.7
    pF.2 = 0.6
    p.D = pD.F2 * pF.2+ p.D.not.F2[i] * (1 - pF.2)
    pF2.D = pD.F2*pF.2/p.D
    p.F2.D.2 <- append(p.F2.D.2,pF2.D)
}

p.F2.D.2
```
```{r}
length(p.F2)
length(p.F2.D.1)
```

## 3

```{r}
par(mfrow = c(1,2))
plot(p.F2, p.F2.D.1, type = "l", ylim = c(0,1))
abline(h = max(p.F2.D.1), col="red", lwd=3, lty=2)
plot(p.D.not.F2, p.F2.D.2, type = "l", ylim = c(0,1))
abline(h = max(p.F2.D.2), col="red", lwd=3, lty=2)
```

## 4
For the p(f2), the increase of p(f2) makes the p(f2|d) increase. In the contrary, for the p(d|not f2), the increase of p(d|not f2) makes p(f2|d) decrease.
It is obvious that p(f2) is worse to be uncertain about, because given 100 simulated numbders, the range of p(f2|d) is greater on p(f2) than p(d|not f2).

## Problem 2

```{r}
d1 <- read.csv(file = "http://faraway.neu.edu/biostats/assn1_dataset1.csv")
d1
```

## 1
H0: There is an association between flu and age(these two variables are not independent)
H1: Not H0

## 2

```{r}
sea_dta <- subset(d1, flu == "seasonal")
swi_dta <- subset(d1, flu == "swine")

par(mfrow = c(2,1))
hist(sea_dta$age, xlim=c(0,90))
hist(swi_dta$age, xlim=c(0,90))
```

## 3
Apparently, these two flu target different groups of people, because the first plot has separate bars and the second plot has centered bars.

## 4

```{r}
cheld_dta <- subset(d1, age < 18 | age > 65)
adt_dta <- subset(d1, 18 <= age & age <= 65)

children_dta <- aggregate(cheld_dta$age, list(cheld_dta$flu), FUN=length)
adult_dta <- aggregate(adt_dta$age, list(adt_dta$flu), FUN=length)
children_dta
adult_dta
```

```{r}
dta_cmb <- children_dta
dta_cmb["adult"] = adult_dta$x
colnames(dta_cmb) <- c("flu","children/elder", "adult")
dta_cmb
```

## 5

```{r}
rownames(dta_cmb) <- dta_cmb$flu
dta_cmb <- data.frame(dta_cmb$`children/elder`, dta_cmb$adult)
dta_cmb
```

```{r}
chisqr <- chisq.test(dta_cmb)
chisqr
```
From the p-value, which is 2.2e-16, we can reject the null hypothesis at any significance level possible. So we can conclude that we have enough evidence to show they have association.
