---
title: "lab7-Chen"
author: "Weixuan Chen"
date: "3/29/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

```{r}
# Download the data
co2=read.table("http://faraway.neu.edu/biostats/lab7_co2.txt", header=TRUE)
temp=read.csv("http://faraway.neu.edu/biostats/lab7_temp.csv")
pop=read.csv("http://faraway.neu.edu/biostats/lab7_population.csv")
```

## Task 1

## 1

```{r}
str(co2)
```

```{r}
str(temp)
```

```{r}
str(pop)
```

```{r}
range(co2$year)
range(temp$Year)
range(pop$year)
```

## 2

```{r}
temp
```


```{r}
# Reshape from wide to long format
temp.long=reshape(temp, varying=list(names(temp)[2:13]), idvar="Year", v.names = "temp", timevar = "Month", direction="long", times = 1:12)
# Change column names
colnames(temp.long)=c("year", "month", "temp")
```

```{r}
temp.long
```

## 3

```{r}
temp.agg=aggregate(temp ~ year, mean, data=temp.long)
co2.agg=aggregate(ppm ~ year, mean, data=co2)
pop.agg=aggregate(popsize ~ year, mean, data=pop)
```

```{r}
temp.agg
co2.agg
pop.agg
```

## 4

```{r}
merged=merge(co2.agg, temp.agg, by="year")
merged=merge(merged, pop.agg, by="year")
```

```{r}
merged
```

## Task 2

## 1

```{r}
par(mfrow=c(1,3))

plot(co2$year, co2$ppm)
plot(temp.long$year, temp.long$temp)
plot(pop$year, pop$popsize)
```

## 2
According to the graph, co2's ppm and pop's popsize variables need transformation. Tranformation can help increase the quality of the dataset because it ensures the data meets the assumptions of the model and thus increase the accuracy of the result. But it also cause the loss of imformation during the transformation.

## 3
H0: there is no relationship between x and y
H1: there is relationship between x and y

## 4

```{r}
temp.lm <- lm(temp ~ year, data = temp.long)
co2.lm <- lm(ppm ~ year, data = co2)
pop.lm <- lm(popsize ~ year, data = pop)
```

```{r}
summary(temp.lm)
```

```{r}
summary(co2.lm)
```

```{r}
summary(pop.lm)
```

## 5

```{r}
# Extract residuals from linear model
temp.resid <- resid(temp.lm)
# TEst the residuals for normality using a Shapiro-Wilk test
shapiro.test(temp.resid)
```

```{r}
# Extract residuals from linear model
co2.resid <- resid(co2.lm)
# TEst the residuals for normality using a Shapiro-Wilk test
shapiro.test(co2.resid)
```

```{r}
# Extract residuals from linear model
pop.resid <- resid(pop.lm)
# TEst the residuals for normality using a Shapiro-Wilk test
shapiro.test(pop.resid)
```
It shows the heteroscedaticity. 

## 6
For the temp~year, we can see the p-values of parameters are significant, which is good indicator of the relationship between y and x. Besides, the R^2 is 0.68, which consider as fine fitting.

For the ppm~year, we can see the p-values are also significant, but R^2 is also too small. This means there is a relationship between y and x but linear regression almost fails to catch this relationship, which is underfitting.

For the popsize~year, we also have the significant p-values, but R^2 is too high. This may mean overfitting issues.

## 7
The assumption is that the distribution of the data will not change.

## 8

```{r}
# New range of explanatory variable year
newyears=data.frame(year=1880:2050)
# Predict response variable based on each regression
pred.temp.vals=predict(temp.lm, newdata=newyears, interval="confidence")
pred.co2.vals=predict(co2.lm, newdata=newyears, interval="confidence")
pred.pop.vals=predict(pop.lm, newdata=newyears, interval="confidence")
```

## 9

```{r}
library(gridExtra)
```

```{r}
library(ggplot2)
```


```{r}
newyears["pred.temp.vals"] = pred.temp.vals[,1]
newyears["pred.co2.vals"] = pred.co2.vals[,1]
newyears["pred.pop.vals"] = pred.pop.vals[,1]

newyears["pred.temp.lwr"] = pred.temp.vals[,2]
newyears["pred.co2.lwr"] = pred.co2.vals[,2]
newyears["pred.pop.lwr"] = pred.pop.vals[,2]

newyears["pred.temp.upr"] = pred.temp.vals[,3]
newyears["pred.co2.upr"] = pred.co2.vals[,3]
newyears["pred.pop.upr"] = pred.pop.vals[,3]

p1 <- ggplot(newyears ,aes(newyears$year,pred.temp.vals))+geom_point()+geom_ribbon(stat="smooth",method="lm",se=TRUE,alpha=0.2)+geom_line(stat="smooth",method="lm") + geom_line(aes(y = pred.temp.lwr), lty = 2) +
  geom_line(aes(y = pred.temp.upr), lty = 2)

p2 <- ggplot(newyears ,aes(newyears$year,pred.co2.vals))+geom_point()+geom_ribbon(stat="smooth",method="lm",se=TRUE,alpha=0.2)+geom_line(stat="smooth",method="lm") + geom_line(aes(y = pred.co2.lwr), lty = 2) +
  geom_line(aes(y = pred.co2.upr), lty = 2)

p3 <- ggplot(newyears ,aes(newyears$year,pred.pop.vals))+geom_point()+geom_ribbon(stat="smooth",method="lm",se=TRUE,alpha=0.2)+geom_line(stat="smooth",method="lm") + geom_line(aes(y = pred.pop.lwr), lty = 2) +
  geom_line(aes(y = pred.pop.upr), lty = 2)

grid.arrange(
  p1,
  p2,
  p3,
  ncol = 3
)
```

## Task 3

## 1

```{r}
cor(merged$ppm, merged$temp)
cor(merged$ppm, merged$popsize)
cor(merged$temp, merged$popsize)
```

```{r}
pairs(merged[,1:3])
```
We can see three variables are positively correlated to each other.

## 2
Since the variables are highly correlated, the change of one variable could cause the change of the other. Therefore, this approach cannot measure the rate precisely.

## 3

```{r}
model <- lm(temp ~ year + popsize + ppm, data = merged)
summary(model)
```
We can see the p-values for each factor is not significant. This indicates we do not have enough evidence to show there is a relationship between y and xs. But we can see R^2 is 0.85, which is pretty good for the model.

## 4
According to the table, it does not have enough evidence to show the causation between temperature and other factors, since we do not reject the null hypothesis for each factor in the summary table above.
